---
author: "Satoshi Kato"
title: iBreakDown plots for LightGBM models
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  html_document:
    fig_caption: yes
    pandoc_args:
      - --from
      - markdown+autolink_bare_uris+tex_math_single_backslash-implicit_figures
    keep_md: yes
    toc: yes
  word_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress = TRUE, 
                     verbose  = TRUE, 
                     root.dir = ".")

knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "", 
                      message = TRUE, 
                      warning = FALSE, 
                      include = TRUE,
                      echo    = TRUE)

set.seed(1)
```

```{r install.requirements, eval = FALSE}
install.packages("DALEX", dependencies = TRUE)
install.packages("r2d3",  dependencies = TRUE)

```

```{r require.packages, message=FALSE}
require(tidyverse)
require(magrittr)
require(xgboost)

require(DALEX)
require(iBreakDown)

```

# Purpose 

Prediction Breakdown according to https://cran.r-project.org/web/packages/iBreakDown/vignettes/vignette_iBreakDown_regression.html


```{r}
library(DALEX)
library(iBreakDown)

```

# Build XGB model

```{r prep.data.train}
head(dragons)

train_data  <- dragons %>% select(-life_length) %>% mutate_all(as.numeric)
train_label <- dragons %>% pull(life_length)
train_xgbDMatrix <- xgb.DMatrix(as.matrix(train_data), label = train_label)

```

For parameter settings, see. https://xgboost.readthedocs.io/en/latest/parameter.html 

```{r parameter.settings}
params <- list(
  booster      = "gbtree",
  objective    = "reg:squarederror",
  eval_metric  = "rmse",
  max_depth = 5,
  colsample_bytree= 0.8,
  subsample = 0.8,
  min_child_weight = 3,
  eta   = 0.05,
  alpha = 0.25,
  gamma = 0
) 

```

```{r xgb.train}
model_xgb <- xgb.train(params  = params, 
                       verbose = 1,
                       data    = train_xgbDMatrix,
                       nrounds = 300)

```

# breakdown obsavation

according to:

https://cran.r-project.org/web/packages/iBreakDown/vignettes/vignette_iBreakDown_classification.html

and:

https://cran.r-project.org/web/packages/iBreakDown/vignettes/vignette_iBreakDown_titanic.html



Following example fail to build explainer.

```{r}
require(DALEX)
explain_xgb <- DALEX::explain(
  model = model_xgb, 
  data = dragons_test,
  y = dragons_test$life_length,
  label = "XGBoost reg:squarederror"
)

```

Thus, several tricks are needed to break down a prediction by XGB model


```{r}
p_fun_xgb <- function(object, newdata) {

  # test.matrix <- as.matrix(newdata)
  newdata <- dplyr::mutate_all(newdata, as.numeric)
  newdata <- xgb.DMatrix(as.matrix(newdata))
  pred <- predict(object, newdata)
  pred
  
}
```

```{r prep.data.test}
test_data  <- dragons_test %>% select(-life_length)
test_label <- dragons_test %>% pull(life_length)

```

```{r}
explain_xgb <- DALEX::explain(
  model = model_xgb, 
  data = test_data,
  y = test_label,
  label = "XGBoost reg/SE",
  predict_function = p_fun_xgb
)

```


# Visualization

HEREAFTER, just Copy & Paste from official vignette:

## local_attributions

```{r}
bd_xgb <- local_attributions(
  explain_xgb,
  new_observation =  test_data[1, ],
  keep_distributions = TRUE)

```


simply print the result.

```{r}
bd_xgb

```

Or plot it.

```{r}
plot(bd_xgb)

```

Use the `baseline` argument to set the origin of plots.

```{r}
plot(bd_xgb, baseline = 0)

```


Use the `plot_distributions` argument to see distributions of partial predictions.

```{r}
plot(bd_xgb, plot_distributions = TRUE)

```

# Calculate uncertainty for variable attributions

```{r}
bdun_xgb <- break_down_uncertainty(
  explain_xgb,
  new_observation =  test_data[1, ],
  path = "average"
)

```

```{r}
plot(bdun_xgb)

```

## Show only top features

```{r fig.height=4, fig.width=4}
# install.packages("r2d3")
require(r2d3)
plotD3(bd_xgb, max_features = 3)

```

